### 2013-04-30 paper/tex/implementation.tex

\section{Implementations}\label{sec:implementation}