### 2013-05-04 essay/main.tex
write essay

\documentclass[11pt, journal, onecolumn]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage[top=1in, bottom=0.75in, left=0.75in, right=0.75in]{geometry}

\begin{document}

\title{The Impact of Cloud Computing in $10$ Years}
\author{Jilong Liao\\jliao2@utk.edu}
\date{\today}
\maketitle

\begin{abstract}
Cloud computing has been very active in both academia and industry field for several years. Major cloud computing vendors, like Amazon, Google and Dropbox, have provided users and businesses many advantages such as reliable and convenient storage, elastic computing power, and simple application deployment, lower IT infrastructure cost, etc. In $10$ years, I will expect cloud computing's influence far beyond the current achievements, but deeply inside mobile technology, massive open online education, online-to-offline business and scientific computing. In this essay, I will discuss this point with my observations and understandings.
\end{abstract}

\input{tex/essay.tex}

\bibliographystyle{cls/model1-num-names}
\bibliography{reference}

\end{document}