### 2013-05-06 essay/main.tex
revise mobile computing and cloud section.

\documentclass[11pt, times]{article}
\usepackage{cite}
\usepackage{url}
%\usepackage{amsmath, amsthm, amsfonts}
\usepackage[top=1in, bottom=0.75in, left=0.75in, right=0.75in]{geometry}

\begin{document}

\title{The Impact of Cloud Computing in $10$ Years}
\author{Jilong Liao\\jliao2@utk.edu}
\maketitle

\begin{abstract}
Cloud computing has been very active in both academia and industry field for several years. Major cloud computing vendors, like Amazon, Google and Dropbox, have provided users and businesses many advantages such as reliable and convenient storage, elastic computing power, and simple application deployment, lower IT infrastructure cost, etc. In $10$ years, I expect the cloud computing's influence is far beyond the current achievements. It should deeply inside mobile technology, education, and local small business. In this essay, I will discuss this point with my observations and understandings.
\end{abstract}

\input{tex/essay.tex}

\bibliographystyle{IEEEtran}
\bibliography{reference}

\end{document}